\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-15 arch.tex
updates

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead and expected latency
  \comment{$\pi \&  \lambda$ both defined for the first time} before
  being able to produce plan to the agent depending on where it is
  located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.45\columnwidth]{figs/TREX_flow}
  \vskip-1cm
  \caption{\small Illustration of state information flow between the
    \texttt{Pilot} and the \texttt{Vehicle} reactors.  Observations
    transit in a bottom-up during synchronization and apply to the
    execution frontier $\tau$.  Goal information flow from the top to
    the bottom and concern the desired future values of state
    variables.\comment{ Fig should be cited in the text}}
  \label{fig:trex:flow}
\end{figure}

Given the above example, the \rx architecture therefore helps abstract
reactor interaction by constraining them to state variables. Each
reactor is therefore agnostic where its {\em external} state
information is managed as long as this information is available and
properly maintained for managing {\em internal} state
information. This {\em internal} state in turn may be used by other
reactors. Such a formalism also constrains reactor interaction to
state information exchange which comes in two forms:

@@

We first show how a reactor can track and identify state when it does
not have any compelling need to deliberate with no future goals. Even
in this condition, \rx requires each reactor to produce at every tick
its {\em internal} state. For those reactors not owning {\em internal}
timelines, \rx makes sure that {\em external} observations are taken
into account for every tick as shown in \cite{py10}. Such a
requirement is enforced to ensure that all reactors share a consistent
view of the world with advancing time.

@@

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal. \comment{Need
  to develop the reason why it is a necessity: primarily to provide
  its internal state to whoever observe it but also simply to ensure
  that its current representation of the world is up to date and still
  consistent}

We introduce this requirement as a new type of flaw for the \eu
framework. This flaw forces the reactor to identify its internal state
for the current tick. By using this flaw one can describe the
synchronization process as:

@@

An internal state flaw can be resolved using one of the following
choices which are evaluated in order:

\item Extend the previous state value to end after this tick (\ie
  restrict its end time to $[\tau+1, \infty)$). \comment{first use of
    $\tau$, so needs to be defined earlier}

All of these choices are evaluated sequentially until a consistent
solution with no flaws at this tick, is identified. In order to do so
we need to make the assumption that the current state value of an
internal timeline does not depend on the future or more accurately
that any choices made during this synchronization will not lead to a
future inconsistency. Such an assumption impacts the set of possible
domains the reactor can support while remaining complete. Take for
example this rule from our \texttt{Shopping} model :

@@

While this model is perfectly acceptable for planning it becomes
problematic when execution is taken into account especially as it puts
a strong emphasis between an evolving plan and an unknown future
(\texttt{Going} to a location and the {\em expected} future outcome of
ending at this location, ties it strongly to the constraint in line
$5$ above).  So for example, assume that the \texttt{Agent.location}
is {\em external} to the reactor which receives the observation
\texttt{Going(Home,SuperMarket)} which synchronization has resolved by
producing the observation \texttt{Go(Home, SuperMarket)} in the
\texttt{Agent} timeline. This would imply that the only possible
outcome on \texttt{Agent.location} is to be \texttt{At(SuperMarket)}
in the directly foreseeable future. In the real world the fact that we
are \texttt{Going} to a certain location does not necessarly imply
that we will succeed on ending up there. In one extreme case due to an
exogenous event, the road to the \texttt{SuperMarket} is closed which
could result in an alternate observation that the robot is
\texttt{At(7th st. \& Main)}. This observation contradicts the
existing plan the reactor is maintaining. Further, it is important to
understand the temporal direction of the \emph{causal link} that is
triggering the inconsistency. At the time the robot observes this
contradicting the \texttt{At} predicate, the rule that generated this
conflict is connected to past observations \texttt{Going} and
\texttt{Go} which cannot be easily retracted without a potential
global impact to the plan implying a huge cost in back-propagation
that would be problematic.

@@

\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}

@@

One final aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. And by doing so we allow for graceful degradation
where other reactors can readapt their plans as they receive this
observation.

\subsubsection{Deliberation : planning for future state evolution}
\label{sec:arch:plan}

The planning process inside a reactor is directly based on the way
\eu work which was described in previous section. Still some simple
alterations were made in order to both include deliberation
information one reactor provide to the architecture -- such as its
deliberation latency and its planning look-ahead both expressed in
ticks -- and the fact that we potentially need to interleave this
deliberation with not the synchronization process but also with the
deliberation of other reactors\footnote{Current implementation of \rx
  is running on a single process and it is the responsibility of the
  agent itself to emulate the multi-threading of the reactors
  deliberation and synchronization}. For this reason while the overal
planning of a \eu reactor strictly relies on the \eu framework
plnanning solver, its slices the execution of Algorithm
\ref{alg:europa:solve} in atomic steps. The interruption is done
around the recursive call (line \ref{li:recurse}) while always
ensuring that the current partial plan mainitained is never proven
inconsistant at any single step. This can be managed by allowing the
call to backtrack until a DecisionPoInt that is not exhausted (ie
$decision.hasNext()$ is \texttt{true}) is find in the call stack.

One critical aspect of this deliberation step is to ensure that at the
exit of this call the $plan$ is not proven inconsistent. Indeed, it is
always possible that the next call made by the agent is the
synchronization which would immediately fail if the $plan$ is
inconsistent before any solving attempt made for synchronization.  For
this reason if the plan is proven inconsistent during a deliberation
step we either backtrack in the decision stack until we find a node fo
the tree that have remaining alternate solution or simply fully relax
the plan in situation there's absolutely no other alternative (by
removing all the goals and keeping only recent observations
produced). These constraints are yet again related to the fact that
for the agent the only crucial functionality of a reactor is to
resolve successfully its synchronization.

The number of flaws between two steps can evolve for the following
reasons that are exogenous to deliberation :

@@

This scope allow to filter out any tokens that are either necessarily
ending before $\tau$ or necessarily starting after. Focusing the
[planning problem to only the tokens that are in a reasonable
future. This window is taken into account by the agent which will
notify the reactor of new {\em internal} goals only when they overlap
this window but also at the deliberation solver in order to reduce the
number of tokens to be evaluated and consequently reduce the cost of
deliberation. As time advance flaws that where initially beyond this
window will become active flaw that the reactor can then evaluate
resulting on an apparent continuous planning of the reactor.

When there are no more flaw is present in the current planning scope a
plan solution is found. This has two effects:

@@

A core aspect of this reactor is that it takes advantage of the
specific sequencing enforced by the \rx architecture along with the
fact that only one $plan$ structure is maintained in the reactor and
shared between synchronization and planning to allow these two
processes to not only be interleaved but impact each other through the
way they manipulate this plan.

The most obvious case is how synchronization allow propagate of {\em
  external} observations as a new tick occur in the plan. In or
presentation of synchronization, we isolated the problem by stating
that we will assume at this point that the reactorhas no future goal
or more accurateely no plan to enforce at this stage. By doing so we
were able to develop the way we augment yhe europpa solver to be able
to propagete {\em external} observations in order to identify the
reactor {\em internal} state for the same tick. By doing so we were
able to focus on synchronization as a state evluation process that can
be resolved within the \eu framework with few extensions. simliarly
we presented deliberation at this fairly high level not necessarily
discussing how the disruptive nature of synchronization in this
process could impact the way this planning process will evolve.

We first anlyse the case during which synchronization occurs after the
planning has vcompleted (ie the last seep of deliberation resulted
into a plan that is considered as complete for the current
look-ahead). As synchronization starts with the plan produced this
will gave a frame for the search process. Therefore this deliberation
will not only be a model-based estimation uniquely based on the {\em
  external} observations but will also include the reactor intentions
as described by the plan produced. Consider for example the Shopping
Agent model described during the \eu presentation and that the
location of the \texttt{agent} is an {\em external} timeline. For
example as, the reactor add the goal to have \texttt{Own} milk and the
current {\em external} states indicated that it was \texttt{At(home)}
it produced the plan in Fig. \ref{fig:shop:exec0}. The {\em external}
part of the plan -- namely the tokens \texttt{Going(home,
  superMarket)} and \texttt{At(SuperMarket)} -- and integrated as
goals for the reactor managing this specific state varaiable {\em
  internally}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}
  \caption{Shopping example : plan produced before synchronization. 
    Tokens in red indicate observations, 
    tokens in blue the goal that this plan attempt to solve. Arrows
    indicate temporal constraints.}
  \label{fig:shop:exec0}
\end{figure}

Now, consider that at the next tick $\tau' = \tau+1$ the reactor managing the
agent location was able to change the location state varaibale to the
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau'$ and lastign for an unlknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization the solver can identify the similarity between this
new observation oand the next token in the existing plan. Which mean
that instead of making a blind exhaustive search of all the possible
implication the synchronization process starts its execution with the
assumption that these new observations are consitent with the current
plan maintained. As a result at the end of synchronization we will
have the plan depicted in Fig. \ref{fig:shop:exec1}. Which just
propagated the information provided by the new observation in the
exisiting plan (restricting in turn the start time of \texttt{Going}
and by propagation of the constraints \texttt{Go} to be {\em exactly}
$\tau'$). While focusing uniquely on the execution frontier ($\tau'$
in Fig. \ref{fig:shop:exec1}), the \eu solver took benefit of the
exisiting plan to evaluate a solution that asumes that the observation
is consistent with what was decided inside this reactor.

>>>>

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as

  each reactor declares its planning look-ahead and expected latency
  \comment{$\pi \& \lambda$ both defined here for the first time}
  before being able to produce plan to the agent depending on where it
  is located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

Given the above example, the \rx architecture therefore helps abstract
reactor interaction by constraining them to state variables. Each
reactor is therefore agnostic where its {\em external} state
information is managed as long as this information is available and
properly maintained for managing {\em internal} state
information. This {\em internal} state in turn may be used by other
reactors. Such a formalism also constrains reactor interaction to
state information exchange which comes in two forms as illustrated in
Figure \ref{fig:trex:flow}:

@@

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal. Even in this
condition, \rx requires each reactor to produce at every tick its {\em
  internal} state. For those reactors not owning {\em internal}
timelines, \rx makes sure that {\em external} observations are taken
into account for every tick as shown in \cite{py10}. Such a
requirement is enforced to ensure that all reactors share a consistent
view of the world with advancing time.

@@

We introduce this requirement as a new type of flaw for the \eu
framework defined as follows:

\begin{definition}
  \label{def:csf}
  A {\em current state flaw} indicates that a reactor {\em internal}
  timeline does not provide a fully defined state value for the
  execution frontier $\tau$. Such a flaw can be result of either of
  the following reasons:

  \begin{itemize}
  \item The absence of a token in this timeline which overlaps the
    instant $\tau$.
  \item An open decision between terminating the last token for this
    timeline or pushing its end in the future.
  \end{itemize}

\end{definition}

The core idea between this flaw is to ensure that at any single tick
the {\em synchronization} solver will identify all the {\em internal}
timelines for which the current state value is yet to be
identified. We handle their resolution based on the following
proposition.

\begin{proposition}
  \label{prop:csf:resolve}
  A {\em current state flaw} can be resolved using one of the
  following choices which are evaluated in the following order:

  \item Extend the previous state value to end after the execution 
    frontier  (\ie restrict its end time to $[\tau+1, \infty)$). %\comment{first use of
    % $\tau$, so needs to be defined earlier}

\end{enumerate}
\end{proposition}

Intuitively the first choice enforces an intertail value assumption
where the reactor will tend to maintain its current as long as the
model and external observations allow it. The second choice favorise
the adavnce of the reactor current plan if any; this direct the {\em
  synchronization} decision toward assuming that its current plan is
properly executed. Finally if none of these two choices did provide a
valid solution -- meaning that both lead to an inconsittant plan --
the reactor tries all the possible state values remaining for this
timeline until it finds one that allows it to identify its state as
required by \rx architecture.

This flaw forces the reactor to identify its internal state
for the current tick. By using this flaw one can describe the
synchronization process as:

@@

\item Resolve the current state value of each internal timelines based
  on Proposition \ref{prop:csf:resolve}.
\end{enumerate}

Steps $2$ \& $3$ are handled by the \eu solver from algorithm
\ref{alg:europa:solve}. We only enforce that the {\em current state
  flaws} are selected last by the algorithm in order to follow the
sequencing given above. As {\em synchronization} focusses on the
execution frontier wee also filter out all flaws that are outside of
the current tick scope. This mean that if this flaw relates to a token
that is {\em necessarily} ending before $\tau$ or starting after
$\tau+1$ then the solver will ignore it. This helps to reduce the
scope of the synchronization to exclusively the current state of the
reactor and reduce dramatically the search space for the {\em
  synchronization} solver. In order to do so we need to make the
assumption that the current state value of an {\em internal} timeline
does not depend on the future or more accurately that any choices made
during this synchronization will not lead to a future
inconsistency. Such an assumption impacts the set of possible domains
the reactor can support while remaining complete. Take for example
this rule from our \texttt{Shopping} model:

@@

While this model is perfectly acceptable for planning it becomes
problematic when execution is taken into account especially since it
puts a strong emphasis between an evolving plan and an unknown future
(\texttt{Going} to a location and the {\em expected} future outcome of
ending at this location, ties it strongly to the constraint in line
$5$ above).  So for example, assume that the \texttt{Agent.location}
is {\em external} to the reactor which receives the observation
\texttt{Going(Home,SuperMarket)} which synchronization has resolved by
producing the observation \texttt{Go(Home, SuperMarket)} in the
\texttt{Agent} timeline. This would imply that the only possible
outcome on \texttt{Agent.location} is to be \texttt{At(SuperMarket)}
in the directly foreseeable future. In the real world the fact that we
are \texttt{Going} to a certain location does not necessarly imply
that we will succeed on ending up there. In one extreme case due to an
exogenous event, the road to the \texttt{SuperMarket} is closed which
could result in an alternate observation that the robot is
\texttt{At(7th st. \& Main)}. This observation contradicts the
existing plan the reactor is maintaining. Further, it is important to
understand the temporal direction of the \emph{causal link} that is
triggering the inconsistency. At the time the robot observes this
contradicting the \texttt{At} predicate, the rule that generated this
conflict is connected to past observations \texttt{Going} and
\texttt{Go} which cannot be easily retracted without a potential
global impact to the plan implying a huge cost in back-propagation
that would be problematic.

@@

% \begin{minipage}[t]{\textwidth}
% \begin{figure}
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}
% \label{lab:model:shopping}
% \caption{\small Model fragment to enforce a \texttt{Going} end condition}
% \end{figure}
% \end{minipage}

@@

One final aspect to take into account is that from a \rx point of
view, \emph{synchronization is the only critical task of a
  reactor}. Such a process is used in order for the agent to identify
its current state and maintain it consistently throughout the agent's
lifetime. For a reactor not being able to find a consistent solution
during synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. And by doing so we allow for graceful degradation
where other reactors can readapt their plans as they receive this
observation.

\subsubsection{Deliberation: Planning for future state evolution}
\label{sec:arch:plan}

Deliberation inside a reactor is based on \eu capabilities described
in Section \ref{sec:europa:arch}. However deliberation within \rx
comes with some augmentation -- deliberation information one reactor
provide to the architecture such as its deliberation latency and its
plan look-ahead both expressed in ticks -- and the fact that we
potentially need to interleave deliberation with not just the
synchronization process but also with deliberation of other
reactors\footnote{The current implementation of \rx is running on a
  single process and it is the responsibility of the agent itself to
  emulate reactor multi-threading for deliberation and
  synchronization.}. For this reason while the overall planning of a
deliberative reactor strictly relies on the \eu framework planning
solver, it slices the execution of Algorithm \ref{alg:europa:solve} in
atomic steps. The interruption is done around the recursive call (line
\ref{li:recurse}) while ensuring that the current partial plan is
never proven inconsistant at any single step. This can be managed by
allowing the call to backtrack until a \textsf{DecisionPoInt} that is
not exhausted (\ie $decision.hasNext()$ is \texttt{true}) is found in
the call stack.

A critical aspect of this deliberation step is to ensure that at the
exit of this call the $plan$ is not inconsistent. % Indeed, it is always
% possible that the next call made by the agent is for synchronization
% which could fail if the $plan$ is inconsistent before any attempt is
% made for synchronization.
If the partial plan is found to be inconsistent during a deliberation
step we either backtrack in the decision stack until a node of the
tree is found with an alternate solution or the plan is relaxed since
there is no other alternative solution. We do this by removing all
goals and keeping only recent observations. % These constraints are yet
% again related to the fact that for the agent the only crucial
% functionality of a reactor is to resolve successfully its
% synchronization.

The number of flaws between two steps \comment{between synch. and
  execution?} can evolve for the following reasons that are exogenous
to deliberation:

@@

\comment{Insert a figure here [AAMAS?] showing $\pi \lambda$} Such a
scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable''
future. This window is taken into account by the \rx agent which will
notify a reactor of new {\em internal} goals not only when they
overlap this reactor specific window but also for the deliberation
solver in order to reduce the number of tokens to be evaluated to
reduce deliberation cost and performance. As time advances flaws that
were initially beyond this temporal horizon will become active which a
reactor can evaluate. This results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

@@

One key aspect of the \rx framework is intertwined plan execution. As
seen above, it is synchronization which drives planning, so the
co-mingling between these two processes is critical to maintaining
state within a robot immersed in a dynamic environment. Also important
is that \rx maintains 
% A core aspect of this reactor is that it takes advantage of the
% specific sequencing enforced by the \rx architecture along with the
% fact that only 
one $plan$ structure in the reactor and shared between synchronization
and planning. This allows these two processes to not only be
interleaved but how they can impact each other via plan manipulation.

The most obvious case is how synchronization allows propagate of {\em
  external} observations when a new tick occurs. In our presentation
of synchronization in Section \ref{sec:arch:synch}, we isolated the
problem by assuming that the reactor has no future goal or more
accurateely no plan to enforce.  By doing so we were able to develop
the way we augment the \eu solver to be able to propagete {\em
  external} observations in order to identify the reactor {\em
  internal} state for the same tick. By doing so we were able to focus
on synchronization as a state evaluation process that can be resolved
within the \eu framework with few extensions. Similarly we presented
deliberation as an abstraction while not detailing how the disruptive
nature of synchronization could impact the way planning could evolve.

We first analyze the case during which synchronization occurs after
planning has completed (\ie the last step of deliberation results in a
plan that is considered complete for the current look-ahead). As
synchronization starts with a plan % produced this will gave a frame for
% the search process. Therefore 
this deliberation will not only be a model-derived estimation uniquely
based on {\em external} observations but will also include reactor
intentions specified in the available partial plan. Consider for
example the Shopping Agent model described where the \texttt{Agent} is
an {\em external} timeline. As the reactor adds the goal to
\texttt{Own} milk and the current {\em external} timeline indicates
that the robot was \texttt{At(home)}, it produces a plan as shown in
Fig. \ref{fig:shop:exec0}. The {\em external} part of the plan --
namely the tokens \texttt{Going(home, superMarket)} and
\texttt{At(SuperMarket)} -- are installed as goals for the reactor
managing this state variable {\em internally}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}
  \caption{\small Shopping example: A plan produced before
    synchronization.  Tokens in red indicate observations, tokens in
    blue the goal that this plan attempts to solve for. Arrows
    indicate temporal constraints.}
  \label{fig:shop:exec0}
\end{figure}

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau'$ and lasting for an unknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization, the solver can identify the similarity between this
new observation and the next token in the existing plan. This implies
that instead of initiating an exhaustive search of all possible
implications, the synchronization process starts its execution with
the assumption that these new observations are consitent with the
currently maintained plan. As a consequence, at the end of
synchronization we will likely have the plan depicted in
Fig. \ref{fig:shop:exec1}. Which just propagated the information
provided by the new observation in the exisiting plan (restricting in
turn the start time of \texttt{Going} and by propagation of the
constraints \texttt{Go} to be {\em exactly} $\tau'$). While focusing
uniquely on the execution frontier ($\tau'$ in
Fig. \ref{fig:shop:exec1}), the \eu solver took benefit of the
exisiting plan to evaluate a solution that asumes that the observation
is consistent with what was decided inside this reactor.
